Dark Control:\\ A Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2017-09-10 paper/paper.tex
Typo + Shorten Venn diagram caption

\begin{figure}[!h]
  \includegraphics[width=1\linewidth]{VennDiagram-2017-09-08.pdf}
  \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}
  This Venn diagram summarizes the relationship between four existing accounts of DMN: in blue colors, the Free Energy Principle \& Predictive Coding, more related to the ”How” DMN works, and in yellow, the Sentinel \& the Semantic hypotheses, more related to "What" DMN does. Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur; here the mental processes and how it relate to observed behavior. For each account, 4 icons describe the core concepts of Time horizon, Neural plasticity, Hierarchy, and Behavioral goal. The "Time horizon" should be understood as a weighting of the expected future outcomes shorter to longer terms into present decision making. Then, "online monitoring" of the environment means detecting and processing salient cues in the sensory environment. The "Predictive Coding" is understood here as the joint minimization of feedforward prediction error and feedback modulation to optimize neural processing; there is thus no notion of an agent with behavioral goal.
  }
  \label{fig:VennDiagram}
\end{figure}

@@

\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
Cortical responses are explained as
emerging from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory processing is constantly calibrated by
top-down modulation from more multi-sensory and associative brain regions
further away from primary sensory cortical regions.
The dynamic interplay between cortical processing levels
may enable learning about aspects of the world by reconciling
gaps between fresh sensory input and expectations computed
based on stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of environmental sensations is
directly compared against the actual environmental input.
A prediction error at one of the processing levels
incurs plasticity changes of neuronal projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In this way,
the predictive coding hypothesis offers explanations for
the constructive, non-deterministic nature of sensory perception
\citep{friston2010free, buzsaki2006rhythms} and
the intimate relation of motor action to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual integration of sensorimotor perception-action cycles
may be maintained by top-down modulation
using a-priori information about the environment.

In short, predictive coding processes
conceptualize updates of the internal representation of the environment
to best accommodate and prepare the organism
for processing the constant influx of sensory stimulation and
performing action on the environment.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account of DMN function based on MDPs.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both functional accounts are plausibilized by
neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections indeed outnumber by far the axonal input projections
existing in the monkey and probably also human brain
\citep{salin1995corticocortical}.
These many and diverse modulatory influences
from higher onto downstream cortical areas
can inject prior knowledge
at every stage of processing environmental information.
%
Moreover,
both accounts provide a parsimonious explanation for why the
human brain's processing load devoted to incoming information decreases
when the environment becomes predictable.
This is because the internal generative
model only requires updates after discrepancies have occurred between
environmental reality and its internally reinstantiated representation.
Increased computation resources are however allocated
when unknown stimuli or
unexpected events are encountered by the organism.
The predictive coding and MDP account hence
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unanticipated environmental challenges.

>>>>

\begin{figure}[!h]
  \includegraphics[width=1\linewidth]{VennDiagram-2017-09-08.pdf}
  \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}
  This Venn diagram summarizes the relationship between four existing accounts of DMN: in blue colors, the Free Energy Principle \& Predictive Coding, more related to the ”How” DMN works, and in yellow, the Sentinel \& the Semantic hypotheses, more related to "What" DMN does. Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur.
  % ; here the mental processes and how it relate to observed behavior. For each account, 4 icons describe the core concepts of Time horizon, Neural plasticity, Hierarchy, and Behavioral goal. The "Time horizon" should be understood as a weighting of the expected future outcomes shorter to longer terms into present decision making. Then, "online monitoring" of the environment means detecting and processing salient cues in the sensory environment. The "Predictive Coding" is understood here as the joint minimization of feedforward prediction error and feedback modulation to optimize neural processing; there is thus no notion of an agent with behavioral goal.
  }
  \label{fig:VennDiagram}
\end{figure}

@@

\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
Cortical responses are explained as
emerging from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory processing is constantly calibrated by
top-down modulation from more multi-sensory and associative brain regions
further away from primary sensory cortical regions.
The dynamic interplay between cortical processing levels
may enable learning about aspects of the world by reconciling
gaps between fresh sensory input and expectations computed
based on stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of environmental sensations is
directly compared against the actual environmental input.
A prediction error at one of the processing levels
induces plasticity changes of neuronal projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In this way,
the predictive coding hypothesis offers explanations for
the constructive, non-deterministic nature of sensory perception
\citep{friston2010free, buzsaki2006rhythms} and
the intimate relation of motor action to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual integration of sensorimotor perception-action cycles
may be maintained by top-down modulation
using a-priori information about the environment.

In short, predictive coding processes
conceptualize updates of the internal representation of the environment
to best accommodate and prepare the organism
for processing the constant influx of sensory stimulation and
performing action on the environment.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account of DMN function based on MDPs.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both functional accounts are supported by
neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections indeed outnumber by far the axonal input projections
existing in the monkey and probably also human brain
\citep{salin1995corticocortical}.
These many and diverse modulatory influences
from higher onto downstream cortical areas
can inject prior knowledge
at every stage of processing environmental information.
%
Moreover,
both accounts provide a parsimonious explanation for why the
human brain's processing load devoted to incoming information decreases
when the environment becomes predictable.
This is because the internal generative
model only requires updates after discrepancies have occurred between
environmental reality and its internally reinstantiated representation.
Increased computation resources are however allocated
when unknown stimuli or
unexpected events are encountered by the organism.
The predictive coding and MDP account hence
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unanticipated environmental challenges.
